Charack: tool for real-time generation of pseudo-infinite virtual worlds for 3D games

### 2009-07-24 paper/paper.tex
- Correcting some mistakes.

\begin{abstract}
In MMO games the player's experience is mainly influencied by the size and details of the virtual world. Technically the bigger the world is, the bigger is the time the player takes to explore all the places. This work presents a tool able to generate pseudo-infinite virtual worlds with different types of terrain. Using a combination of algorithms and content management methods, the tool is able to create beachs, islands, bays and cost lines that imitates the real world landscapes. One of the contributions of the tool is the ability to generate giant pieces of land focusing on the coast line generation. The development of the present work aimed to handle separately the generation of content for all elements in the world (continents, terrains, etc.). The main point in the work is the coast line generation, not the content inside the continents.
\end{abstract}

@@

The computer games market has been evolving considerably over the years. Since the first console, the hardware performance has increased and new graphic technologies were developed, resulting in a wide range of themes and game styles. In the multiplayer games, players interact with other human beings and also with NPCs, which are represented by virtual characters. The kind of game is popular and the social interaction between players is a matter of research ~\cite{onlineGaming,ducheneaut}. In the category of multiplayer games there are the massively multiplayer online (MMO) ones, which are online games featuring large number of players interacting with each other in a persistent virtual world.

@@

To ensure acceptable use of memory and CPU, only the elements inside the user's view are generated. As the user moves through the city, new cells are added to the user's view and the new content is generated. Whe the cell leaves the user's view, it is removed from memory and its resources are released. The cells are placed in square loops around the user and are considered inside the user's view if it is close enough and if it is located within a 120° view range.

@@

One of the texturization methods presented are the use of a large image able to cover the whole world. Although this approach is useful for small scenes, it is not suitable for large scenes nor virtual worlds, since the image size could reach prohibitive proportions. To avoid this problem, an reticulate approach is suggested ~\cite{CSHD03}; using that approacha a texture cell is created so that the a combination of them produces a smooth and continuous and textured plane. The results are acceptable, but it is not visually attractive to the user, because the landscape features an unusual pattern that will never exist in the real world. In order to achieve a better visual result, an approach proposed by ~\cite{LN03} is suggested. In that approach, a set of pre-defined border textures is combined with reticulate textures. The idea is to apply the border textures on an already textured plane, but using transparency in the border textures. As a consequene, border textures will overlap the some reagions of the plane and produce a less homogenous landscape, which creates a better visual result.

@@

The use of fractals and their combinations with other methods are also used to proceduraly generate planets ~\cite{LindaOndrej2007}. In this approach, the generated world is not infinite, but it is spherical and simulates the view of the planet Earth. Starting with a recursive subdivision of an octahedron, a spherical world is created and it is used as a first step to sevel algorithms of heighmap generation. The heightmap is obtained as a result of several techniques, each one featuringits peculiarities and results. Among the techniques used are: generation using random failures, midpoint displacement (including a multifractal variation) and Perlin noise (and its many variations).

@@

Another approach for procedural terrain generation in real time uses fractals affected by erosion ~\cite{terrain_generation}. As pointed out, the increasing processing power of home computers has enabled games to make use erosion simulations almost in real time. As a consequence, an approach envolving fractals and erosion is possible to be reached, resulting in more realistic terrain. In order to achieve that, a heightmap was created from a Voronoi diagram, where each vertices of the diagram is called {\it functionality point}; the value of each cell in the heightmap is obtained through a linear combination of distances of the closest functionality points. Once the heightmap is generated, it is combined with a noise algorithm in order to produce mountains. The sharp lines introduced by the well-defined Voronoi diagram are removed with the application of a turbulence filter. For the erosion simulation an hybrid method is used, which is a combination of two techniques: the thermal ~\cite{MusgraveErosao} and the hydraulic erosion.

@@

The procedurally generated world approach ~\cite{LindaOndrej2007} is very close to the concept of content generation aimed for Charack. In that work, a spherical planet is created as a result of a recursive division of a geometric shape, then noise functions are applied to the mesh to generate the heightmap. There is no distinction between the content generation approache for continents and the content generation for the lands within the continents. As a result the continents are created by flooding the heightmap with a water plane, which will produce the coast lines based on height of the sea level and the amount of ripples in the topography. The content itself is not generated on demand. Charack was created from an evolution of this idea, but subject to some limitations. The world created by Charack handles differently the generation of continents, coast lines and the heighmap within them. Each of these generation processes has its own peculiarities, which can be tweaked in order to customize specific details in the world with no interference in the other processes. This is the case of the creation of long beach on the coast without changing the basic structure or the topography of the continent that the beach is inserted. Charack also generates the content on demand, however it doesn't use a spherical approach, the resulting world contained within a plan.

The generation of a virtual world as a result of recursive subdivisions of a quadtree ~\cite{dollins-thesis} is very similar to the Charack proposal. In that work, a world with huge proportions is created and its content is generated on demand as the user moves. The heighmap is created in a parameterized and multi-resolution way, so the closer the user of place, the greater is the amount of detail there. There is also no distinction in the generation process of continents/coastlines/land content. The heightmap generation proposed is used by Charack, however continents and the coastlines generation process are completely different.

@@

\begin{figure}
\centering
\resizebox{6cm}{!}{\includegraphics{images/english/planejamento_estrutura_basica.png}}
\caption{Charack's basic structure}
\label{fig:planejamento_estrutura_basica}
\end{figure}

@@

The resulting virtual world technically has height and width defined by the maximum size of a signed integer. It is physically impossible to generate a MM with such proportions. Since the MM is smaller than the virtual world, an MM's entry {\tt (i, j)} represents sevel pixels in the virtual world. Figure \ref{fig:mapeamento_macro} illustrates the MM mapping process.

@@

As the figure ilustrates the smaller is the MM, the more pixels in the virtual world will be represented by the same entry in the MM. If the virtual world had {\tt 1000x1000} as its size and the MM had {\tt 10x10} as its size, for instance, it means that for each pixel of the MM represents 100 pixels in the virtual world. If one of these pixels is described in the MM as coast, then there is an area of {\tt 100x100} pixels in the virtual world that must be a coast. To solve this problem the CG acts. When any of those pixels in that particular area of the virtual world is analyzed by the CG, it will work on it and return it with different values, which will result in a coast line for that area, not am area entirely filled with land or water.

@@

The solution makes the player able to look at the screen and see a slice of the virtual world, but with no explicit divisions in the world, such as cells. Unlike what was done in ~\cite{infinicity} where the user's view is a cone, the Charack user's view is a square centered on the user. Using the user position {\tt (x, y, z)}, the visible content arround that coordinate is sliced from the world and drawn in screen. If the player reaches a place or a complete slice is not possible, such as the world boundaries, no content is displayed beyond the world area.

In order to texture the sliced data a set of images are interpolated and managed by the shading language GLSL ~\cite{glsl}. The height value of the pixel defines the interpolation weight of each texture. As a consequence, a sand texture has highter weight for pixels featuring a low height value, for instance. Figure ~\ref{fig:set_texturas} shows all the available textures.

@@

The main idea for the terrain height generation is the use a parametric function that informs the height value of each vertex. The function is seeded with the point location in the world. As a consequence, the function is able to describe all the height information in the world with no limitions concerning the world size. The processing time is related to the size of the user's view, not related to the world size, since the function uses the point information to calculate its height value. The height values are generated with a Perlin noise function ~\cite{Perlin85}.

@@

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/torben.png}}
\caption{Random planets generated with different seeds. (a) and (b) maps featuring height value information; (c) map featuring only information about what is wahter/ocean.}
\label{fig:torben}
\end{figure}
%~\cite{torben}

@@

This figure ilustrates a place in the virtual world that represents the transition between two different points of the MM (a land point and a water point). To explain what is happening, assume the tool is drawing the world at position {\tt (x, y, z)}, which is the mapping result of a pixel {\tt (i, j)} in the MM, which is described as land; as the tool increases the coordinate in order to draw the landscape, each new position is mapped to the MM. If the result of the mapping process of new coordinate, {\tt (x + 1, y, z))} for instance, is still the point {\tt (i, j)} in the MM, then the tool will again draw a land pixel on the screen. Assuming that only at point {\tt (x + 10, y, z)} the pixels start being mapped to a different pixel in the MM, such as {\tt (i + 1, j)} (and {\tt (i + 1, j)} is described as land), then all points before the position {\tt (x + 10, y, z)} are drawn as water and all pixels after that location are drawn as land.

The figure shows clearly when the world coordinates start beeing mapped to a different entry in the MM, which is when the tool replaces the lang rendering with water rendering. As a consequence of no algorithm being applied to generate content for that transition area, the user will move along the coastline and will see only straight lines.

@@

The mapping process of the pixels of the virtual world to the MM produces very unrealistic landscapes. A real world beach has a natural curvature and hardly have a length of 20km in a perfectly straight configuration, as the beaches generated by Charack. Although the objective of this work is not to create photo-realistic landscapes, such unreal beaches are not acceptable. To circumvent this problem, a coastline disturbance is applied to the locations where the mapping process is made between two MM points, one of them described as land and the other one described as water. The algorithm is described below.

@@

From that moment, the first step of the coastline disturbance algorithm is performed. Using the current MM as its input, the algorithm scan each MM's pixel, updating the descriptor of pixels that represent a coast. A pixel is said to be coast when at least one of its neighbors is water. After the algorithm end, the MM contains the three types of terrain described before (water, land and offshore). The next step to apply disturbance to the coastline is the content generation based on the descriptor of each pixel in the MM. When the tool is creating content to draw on the screen, each pixel being drawn is tested against its descriptir in the MM. If the pixel is mapped to a land pixel in the MM, then the function will set a height value for that point. If the pixel is mapped to a water pixel in the MM, then the function will set the sea level height to that point. Finally if the pixel is mapped to a offshore pixel in the MM, then the function will disturb the land/water information of that mapping process, which will result on a non-straight coast line. Figure ~\ref{fig:funcionamento_quebra_linearidade} illustrates the algorithm.

\begin{figure}
\centering
\resizebox{7cm}{!}{\includegraphics{images/english/funcionamento_quebra_linearidade.png}}
\caption{Coastline disturbance algorithm}
\label{fig:funcionamento_quebra_linearidade}
\end{figure}

@@

The implementation of that process is based on a noise function and random numbers with a parametric function deciding what is land and what is water for all pixels described as a coast in the MM. Using the pixel position in the block {\tt RA}, the function maps that information into a spectrum of values created by a Perlin noise function. What the parametric function does is check if the hash of the pixel being analyzed is inside or outside of the spectrum. The process can be imagined as a height test against a small heightmap (which is created as result of the noise spectrum): if the return of the noise function for that is greater than a certain value (which is the granularity of the block being analyzed), then it is classified as land, otherwise it is classified as water. The higher is the granularity of the block, the greater is the amount of land on that location. Figure ~\ref{fig:plot_ruido_quebra_linearidade} illustrates the small heightmap generated by the costline disturbance algorithm when block {\tt RA} is being processed.

\begin{figure}
\centering
\resizebox{2cm}{!}{\includegraphics{images/plot_ruido_quebra_linearidade.png}}
\caption{Small heightmap generated by the costline disturbance algorithm}
\label{fig:plot_ruido_quebra_linearidade}
\end{figure}

@@

The coastline disturbance algorithm minimizes the problem of unrealistic continents lines, but the outcome is not quite good enought. When Charack is rendering a slice of the world, for each pixel described as a land a height vaule is set to the pixel; the same applies to the pixels that are described as water, but in that case the height value is always the same (the sea level). As a direct result of that approach if the tool is drawing a set of pixels describing a mountan and the next pixels are described as water, the landscape will features a "step". It happens because the mountain chain was generated very close to the water, which means that its rendering is abruptly interrupted when Charack finds pixels described as water. Although there are cliffs in the real world, they are not present in all coasts. To solve this problem, a special algorithm is applied in order to create beachs in certain locations of the world, which makes the generated landscape looks more realistic.

@@

This section aims to evaluate each of the techniques used in the content generation process, explaining the obtained results for each approach. It is important to highlight that Charack's purpose is not the generation of real or photo-realistic content, but elements that can be used to create a 3D game scene. A result is classified as graphically acceptable if it can be integrated into a game and not surprive player in a negative way, such as a pyramidal mountain instead of a smooth mountain.

@@

The coastline generation is composed of two main elements, a global and a local one. The global one only uses only data available in the MM in order to create the coastlines, as described in section ~ \ref{sec:visaomicromacro}. The final result for that approach is aunreal straight coastline. Figure ~\ref{fig:charack_final_costa_reta} shows two completely straight coastlines.

@@

Figures ~\ref{fig:charack_final_costa_3} and ~\ref{fig:charack_final_costa_4} show final result obtained with the combination of all the previously described algorithms: coastline dirturbance, beach disturber and island generator.

@@

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/charack_final_costa_4.png}}
\caption{Coastline featuring beachs with different sizes}
\label{fig:charack_final_costa_4}
\end{figure}

>>>>

\begin{abstract}
In MMO games the player's experience is mainly influenced by the size and details of the virtual world. Technically the bigger the world is, the bigger is the time the player takes to explore all the places. This work presents a tool able to generate pseudo-infinite virtual worlds with different types of terrain. Using a combination of algorithms and content management methods, the tool is able to create beaches, islands, bays and cost lines that imitates the real world landscapes. One of the contributions of the tool is the ability to generate giant pieces of land focusing on the coast line generation. The development of the present work aimed to handle separately the generation of content for all elements in the world (continents, terrains, etc.). The main point in the work is the coast line generation, not the content inside the continents.
\end{abstract}

@@

The computer games market has been evolving considerably over the years. Since the first console, the hardware performance has increased and new graphic technologies were developed, resulting in a wide range of themes and game styles. In the multiplayer games, players interact with other human beings and also with NPCs, which are represented by virtual characters. The kind of game is popular and the social interaction between players is a matter of research ~\cite{onlineGaming,ducheneaut}. In the category of multiplayer games there are the massively multiplayer on-line (MMO) ones, which are on-line games featuring large number of players interacting with each other in a persistent virtual world.

@@

To ensure acceptable use of memory and CPU, only the elements inside the user's view are generated. As the user moves through the city, new cells are added to the user's view and the new content is generated. When the cell leaves the user's view, it is removed from memory and its resources are released. The cells are placed in square loops around the user and are considered inside the user's view if it is close enough and if it is located within a 120° view range.

@@

One of the texturization methods presented are the use of a large image able to cover the whole world. Although this approach is useful for small scenes, it is not suitable for large scenes nor virtual worlds, since the image size could reach prohibitive proportions. To avoid this problem, an reticulate approach is suggested ~\cite{CSHD03}; using that approach a texture cell is created so that the a combination of them produces a smooth and continuous and textured plane. The results are acceptable, but it is not visually attractive to the user, because the landscape features an unusual pattern that will never exist in the real world. In order to achieve a better visual result, an approach proposed by ~\cite{LN03} is suggested. In that approach, a set of pre-defined border textures is combined with reticulate textures. The idea is to apply the border textures on an already textured plane, but using transparency in the border textures. As a consequence, border textures will overlap the some regions of the plane and produce a less homogeneous landscape, which creates a better visual result.

@@

The use of fractals and their combinations with other methods are also used to procedurally generate planets ~\cite{LindaOndrej2007}. In this approach, the generated world is not infinite, but it is spherical and simulates the view of the planet Earth. Starting with a recursive subdivision of an octahedron, a spherical world is created and it is used as a first step to several algorithms of heightmap generation. The heightmap is obtained as a result of several techniques, each one featuring its peculiarities and results. Among the techniques used are: generation using random failures, midpoint displacement (including a multifractal variation) and Perlin noise (and its many variations).

@@

Another approach for procedural terrain generation in real time uses fractals affected by erosion ~\cite{terrain_generation}. As pointed out, the increasing processing power of home computers has enabled games to make use erosion simulations almost in real time. As a consequence, an approach involving fractals and erosion is possible to be reached, resulting in more realistic terrain. In order to achieve that, a heightmap was created from a Voronoi diagram, where each vertex of the diagram is called {\it functionality point}; the value of each cell in the heightmap is obtained through a linear combination of distances of the closest functionality points. Once the heightmap is generated, it is combined with a noise algorithm in order to produce mountains. The sharp lines introduced by the well-defined Voronoi diagram are removed with the application of a turbulence filter. For the erosion simulation an hybrid method is used, which is a combination of two techniques: the thermal ~\cite{MusgraveErosao} and the hydraulic erosion.

@@

The procedurally generated world approach ~\cite{LindaOndrej2007} is very close to the concept of content generation aimed for Charack. In that work, a spherical planet is created as a result of a recursive division of a geometric shape, then noise functions are applied to the mesh to generate the heightmap. There is no distinction between the content generation approach for continents and the content generation for the lands within the continents. As a result the continents are created by flooding the heightmap with a water plane, which will produce the coast lines based on height of the sea level and the amount of ripples in the topography. The content itself is not generated on demand. Charack was created from an evolution of this idea, but subject to some limitations. The world created by Charack handles differently the generation of continents, coast lines and the heightmap within them. Each of these generation processes has its own peculiarities, which can be tweaked in order to customize specific details in the world with no interference in the other processes. This is the case of the creation of long beach on the coast without changing the basic structure or the topography of the continent that the beach is inserted. Charack also generates the content on demand, however it does not use a spherical approach, the resulting world contained within a plan.

The generation of a virtual world as a result of recursive subdivisions of a quadtree ~\cite{dollins-thesis} is very similar to the Charack proposal. In that work, a world with huge proportions is created and its content is generated on demand as the user moves. The heightmap is created in a parameterized and multi-resolution way, so the closer the user of place, the greater is the amount of detail there. There is also no distinction in the generation process of continents/coastlines/land content. The heightmap generation proposed is used by Charack, however continents and the coastlines generation process are completely different.

@@

\begin{figure}
\centering
\resizebox{5.5cm}{!}{\includegraphics{images/english/planejamento_estrutura_basica.png}}
\caption{Charack's basic structure}
\label{fig:planejamento_estrutura_basica}
\end{figure}

@@

The resulting virtual world technically has height and width defined by the maximum size of a signed integer. It is physically impossible to generate a MM with such proportions. Since the MM is smaller than the virtual world, an MM's entry {\tt (i, j)} represents several pixels in the virtual world. Figure \ref{fig:mapeamento_macro} illustrates the MM mapping process.

@@

The smaller is the MM size, the more pixels in the virtual world will be represented by the same entry in the MM. If the virtual world had {\tt 1000x1000} as its size and the MM had {\tt 10x10} as its size, for instance, it means that for each pixel of the MM represents 100 pixels in the virtual world. If one of these pixels is described in the MM as coast, then there is an area of {\tt 100x100} pixels in the virtual world that must be a coast. To solve this problem the CG acts. When any of those pixels in that particular area of the virtual world is analyzed by the CG, it will work on it and return it with different values, which will result in a coast line for that area, not am area entirely filled with land or water.

@@

The solution makes the player able to look at the screen and see a slice of the virtual world, but with no explicit divisions in the world, such as cells. Unlike what was done in ~\cite{infinicity} where the user's view is a cone, the Charack user's view is a square centered on the user. Using the user position {\tt (x, y, z)}, the visible content around that coordinate is sliced from the world and drawn in screen. If the player reaches a place or a complete slice is not possible, such as the world boundaries, no content is displayed beyond the world area.

In order to texture the sliced data a set of images are interpolated and managed by the shading language GLSL ~\cite{glsl}. The height value of the pixel defines the interpolation weight of each texture. As a consequence, a sand texture has higher weight for pixels featuring a low height value, for instance. Figure ~\ref{fig:set_texturas} shows all the available textures.

@@

The main idea for the terrain height generation is the use a parametric function that informs the height value of each vertex. The function is seeded with the point location in the world. As a consequence, the function is able to describe all the height information in the world with no limitation concerning the world size. The processing time is related to the size of the user's view, not related to the world size, since the function uses the point information to calculate its height value. The height values are generated with a Perlin noise function ~\cite{Perlin85}.

@@

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/torben.png}}
\caption{Random planets generated with different seeds. (a) and (b) maps featuring height value information; (c) map featuring only information about what is water/ocean.}
\label{fig:torben}
\end{figure}
%~\cite{torben}

@@

This figure illustrates a place in the virtual world that represents the transition between two different points of the MM (a land point and a water point). To explain what is happening, assume the tool is drawing the world at position {\tt (x, y, z)}, which is the mapping result of a pixel {\tt (i, j)} in the MM, which is described as land; as the tool increases the coordinate in order to draw the landscape, each new position is mapped to the MM. If the result of the mapping process of new coordinate, {\tt (x + 1, y, z))} for instance, is still the point {\tt (i, j)} in the MM, then the tool will again draw a land pixel on the screen. Assuming that only at point {\tt (x + 10, y, z)} the pixels start being mapped to a different pixel in the MM, such as {\tt (i + 1, j)} (and {\tt (i + 1, j)} is described as land), then all points before the position {\tt (x + 10, y, z)} are drawn as water and all pixels after that location are drawn as land.

The figure shows clearly when the world coordinates start being mapped to a different entry in the MM, which is when the tool replaces the land rendering with water rendering. As a consequence of no algorithm being applied to generate content for that transition area, the user will move along the coastline and will see only straight lines.

@@

The mapping process of the pixels of the virtual world to the MM produces very unrealistic landscapes. A real world beach has a natural curvature and hardly have a length of 20 Km in a perfectly straight configuration, as the beaches generated by Charack. Although the objective of this work is not to create photo-realistic landscapes, such unreal beaches are not acceptable. To circumvent this problem, a coastline disturbance is applied to the locations where the mapping process is made between two MM points, one of them described as land and the other one described as water. The algorithm is described below.

@@

From that moment, the first step of the coastline disturbance algorithm is performed. Using the current MM as its input, the algorithm scan each MM's pixel, updating the descriptor of pixels that represent a coast. A pixel is said to be coast when at least one of its neighbors is water. After the algorithm end, the MM contains the three types of terrain described before (water, land and offshore). The next step to apply disturbance to the coastline is the content generation based on the descriptor of each pixel in the MM. When the tool is creating content to draw on the screen, each pixel being drawn is tested against its descriptor in the MM. If the pixel is mapped to a land pixel in the MM, then the function will set a height value for that point. If the pixel is mapped to a water pixel in the MM, then the function will set the sea level height to that point. Finally if the pixel is mapped to a offshore pixel in the MM, then the function will disturb the land/water information of that mapping process, which will result on a non-straight coast line. Figure ~\ref{fig:funcionamento_quebra_linearidade} illustrates the algorithm.

\begin{figure}
\centering
\resizebox{6.6cm}{!}{\includegraphics{images/english/funcionamento_quebra_linearidade.png}}
\caption{Coastline disturbance algorithm}
\label{fig:funcionamento_quebra_linearidade}
\end{figure}

@@

The implementation of that process is based on a noise function and random numbers with a parametric function deciding what is land and what is water for all pixels described as a coast in the MM. Using the pixel position in the block {\tt RA}, the function maps that information into a spectrum of values created by a Perlin noise function. What the parametric function does is check if the hash of the pixel being analyzed is inside or outside of the spectrum. The process can be imagined as a height test against a small heightmap (which is created as result of the noise spectrum): if the return of the noise function for that is greater than a certain value (which is the granularity of the block being analyzed), then it is classified as land, otherwise it is classified as water. The higher is the granularity of the block, the greater is the amount of land on that location. Figure ~\ref{fig:plot_ruido_quebra_linearidade} illustrates the small heightmap generated by the coastline disturbance algorithm when block {\tt RA} is being processed.

\begin{figure}
\centering
\resizebox{2cm}{!}{\includegraphics{images/plot_ruido_quebra_linearidade.png}}
\caption{Small heightmap generated by the coastline disturbance algorithm}
\label{fig:plot_ruido_quebra_linearidade}
\end{figure}

@@

The coastline disturbance algorithm minimizes the problem of unrealistic continents lines, but the outcome is not quite good enough. When Charack is rendering a slice of the world, for each pixel described as a land a height value is set to the pixel; the same applies to the pixels that are described as water, but in that case the height value is always the same (the sea level). As a direct result of that approach if the tool is drawing a set of pixels describing a mountain and the next pixels are described as water, the landscape will features a "step". It happens because the mountain chain was generated very close to the water, which means that its rendering is abruptly interrupted when Charack finds pixels described as water. Although there are cliffs in the real world, they are not present in all coasts. To solve this problem, a special algorithm is applied in order to create beaches in certain locations of the world, which makes the generated landscape looks more realistic.

@@

This section aims to evaluate each of the techniques used in the content generation process, explaining the obtained results for each approach. It is important to highlight that Charack's purpose is not the generation of real or photo-realistic content, but elements that can be used to create a 3D game scene. A result is classified as graphically acceptable if it can be integrated into a game and not surprise the player in a negative way, such as a pyramidal mountain instead of a smooth mountain.

@@

The coastline generation is composed of two main elements, a global and a local one. The global one only uses only data available in the MM in order to create the coastlines, as described in section ~ \ref{sec:visaomicromacro}. The final result for that approach is a unreal straight coastline. Figure ~\ref{fig:charack_final_costa_reta} shows two completely straight coastlines.

@@

Figures ~\ref{fig:charack_final_costa_3} and ~\ref{fig:charack_final_costa_4} show final result obtained with the combination of all the previously described algorithms: coastline disturbance, beach disturber and island generator.

@@

\begin{figure}
\centering
\resizebox{8cm}{!}{\includegraphics{images/charack_final_costa_4.png}}
\caption{Coastline featuring beaches with different sizes}
\label{fig:charack_final_costa_4}
\end{figure}
